---
title: "HW02 Markdown Reprt"
author: "Neha Patwardhan, Joy Machado"
date: "January 31, 2016"
output: html_document
---

###Markdown Report

###Required Libraries to run the R Script
```{r echo=FALSE}
require(dplyr)
require(ggplot2)
require(markdown)
```

### R Script
```{r }
# Read the data from the Output file of the Map Reduce Program as an input for the R Script
args <- commandArgs(TRUE)
# The path of the output is the command line argument for the R script
srcFile <- args[1]
files <- dir(srcFile, pattern='part-r-', full.names = TRUE)


outputs <- lapply(files, read.csv, header=FALSE, sep="\t")
output <- bind_rows(outputs)
colnames(output) <- c("carrier","price","month","frequency")

# Sort the input to the R scipt in the descending order of Frequency
sorted <- output[order(-output$frequency),]

# Rank the outputsin order of Frequency after it has been sorted.
# The Airlines with the highest frequency will have the highest Rank
sorted$rank <- rank(-sorted$frequency,ties.method="min")

# Assign Dense Ranking to find the real rank of the Airlines as a group 
sorted$drank <- rep(1:length(rle(sorted$rank)$values),rle(sorted$rank)$lengths)

# Keep only the top 10 airlines with the highest frequency
sortedNew <- sorted[sorted$drank <= 10,]




```


###Plotting the graph of the Average Price of Airlines vs Month
```{r, echo=FALSE}
plot1 <- ggplot(data=sortedNew,aes(x=sortedNew$month,y=sortedNew$price,color=sortedNew$carrier)) + geom_line() + scale_color_manual(name="Top 10 Airlines",values = c("red","green","blue","orange","purple","pink","yellow","black","brown","maroon")) + ggtitle("Average price of airlines per month") + labs(x="Month(1-12)",y="Average Price(USD)")

plot1
```

# Explanation of our Map Reduce Program
  
# Mapper
`The Mapper reads the files from the input and goes through the sanity check
`If a record passes through the sanity check, we have kept a lookout for the carriers active in 2015. If the airlines is active in 2015, we add it to a    Global set.
`The carrier name is set as the Key of the mapper.
`A flight object is created with the year, month and average price of each record and provided to the Reducer as the value.
  
#Reducer
`The Reducer accepts the Carrier name as the Key and we extract the month, and averge price of each flight.
`The average price is computed and the frequency for each key is calcualted.
`If a flight is present in the Global set, only then we print it to the Output files of the Program.
  
  
# Input for R
`The output file of the map reduce program is the input file for the R script.
`The R script sorts the output in the descending manner of Frequency.
`A rank is assigned to each record.
`A Dense rank is then assigned to Ranked and Sorted data set.
`Only the Airlines with the top 10 frequencies are considered.
`This is then given as an input to graph which plot the Average Price of each airlines against each month.

   
# Requirements
`Install R 3.2.3
`Run the markdown file in R Studio
`Install the following packages in R
` dplyr
`, ggplot2
`, markdown

`Also have pandoc installed on your system to view the markdown as a HTML




   
   
  

